Bayesian multivariate skew-normal finite mixture model for analysis of infant development trajectories

### 2019-07-02 manuscript/manuscript.tex
edits

\begin{abstract}
In studies of infant motor development, a crucial research goal is to identify latent clusters of infants that experience delayed development, as this is a known risk factor for adverse outcomes later in life. However, there are a number of statistical challenges in modeling infant development: the data are typically skewed, exhibit intermittent missingness, and are highly correlated across the repeated measurements collected during infancy. Using data from the Nurture study, a cohort of over 600 mother-infant pairs followed from pregnancy to 12 months postpartum, we develop a flexible Bayesian latent class model for the analysis infant motor development. Our model has a number of attractive features. First, we adopt the multivariate skew normal distribution with cluster-specific parameters that accommodate the inherent correlation and skewness in the data. Second, we model the cluster membership probabilities using a novel application of the P\'olya-Gamma data-augmentation scheme, thereby improving predictions of the cluster membership allocations. Lastly, we impute missing responses under missing at random assumption by drawing from appropriate conditional multivariate skew  normal distributions. Bayesian inference is achieved through straightforward Gibbs sampling, and can be implemented in available software such as \texttt{R}.  Through simulation studies, we show that the proposed model yields improved inferences over models that ignore skewness. In addition, our imputation method yields improvements compared to conventional missing data methods, including multiple imputation and complete or available case analysis. When applied to Nurture data, we identified two distinct development clusters: one characterized by delayed “U-shaped” development and a higher percentage of male infants and another characterized by more steady development and a lower percentage of males. The clusters also differed in terms of key demographic variables, such as infant race and maternal pre-pregnancy body mass index. These findings can aid investigators in targeting interventions during this critical early-life developmental window.
\end{abstract}

@@

Mixtures of multivariate non-symmetric distributions such as the multivariate skew-normal (MSN) distribution allow for the nuances of the marginal density to be captured with a more parsimonious set of mixture components. Mixtures of MSN distributions have been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010 and others), however in these models, focus lies primary on marginal density estimation, and inference on the mixture components (i.e. clusters) is often not discussed. More recently, the mixtures of skew-$t$ factor analysis (MSTFA) model has been proposed for settings in which cluster-specific inference is of primary interest (Lin \textit{et al.} 2018). However, an important feature not included in the MSTFA is the ability to explain individual-level cluster membership as a function of covariates of interest. Additionally, parameter estimation proposed by Lin et al. for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, namely the ability to incorporate prior information into a model and make posterior probability statements. Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya-Gamma distribution.

@@

A ubiquitous feature of repeated measures studies is loss of data due to intermittent missingness and attrition. In the Bayesian setting, the standard approach to dealing with missing data is to perform multiple imputation, whereby $m$ imputed data sets are generated from a specified imputation model. After $m$ complete data sets are obtained, parameter estimates are combined across each data set to produce a final set of parameter estimates (Gelman \textit{et al.} 2013). This approach is not only computationally burdensome, requiring storage and analysis of an $m \times n_{rows} \times n_{cols}$ data array in addition to multiplication of total model run time by a factor of $m$, but it has been shown to produce unreliable inferences (Zhou and Reiter, 2010). We instead include an ``online" imputation imputation step in our Gibbs sampling procedure, whereby missing outcomes are updated at each iteration. This approach greatly increases the number of opportunities for exploration of the missing data parameter space.

@@

The analysis of infant development trends in the Nurture data presents a number of statistical challenges that motivate our proposed model. First, as depicted in Figure \ref{fig:skew_resids}, the residuals from repeated measures models of Bayley composite scores exhibit skewness even after adjusting for covariates such as race, sex, and birthweight. This suggests that the assumption of conditional normality made by standard repeated measures models is violated, and a distinguishing feature of the data, namely skewness, is not being accounted for.

@@

The Nurture data also feature intermittent missingness in Bayley composite scores throughout the study period. Of the total cohort (N = 666), 429 (64.4 \%) observations were available at three months, 435 (65.3 \%) observations were available at six months, 418 (62.8 \%) observations were available at nine months, and 437 (65.6 \%) observations were available at twelve months.

@@

A primary goal of the Nurture study is to identify clusters of infants characterized by distinct motor development trajectories. To address this aim, we propose a flexible finite mixture model that accommodates relevant features of the data, such as skewness and dependence among the responses. For $i = 1,...,n$, let $\mathbf{y}_{i}=(y_{i1},\ldots,y_{iJ})^T$ be a $J \times 1$ vector of responses (i.e., Baley scores) for subject $i$ across the $J$ measurement occasions. For the analysis of the Nurture data, we propose a finite mixture model of the form
\begin{equation}
f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ki} f(\mathbf{y}_i|\boldsymbol\theta_k),
\end{equation}
where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$) and $\pi_{ki}$ is a subject-specific mixing weight representing the probability that subject $i$ belongs to cluster $k$. For now we assume that $K$ is fixed; in Section 4, we discuss model selection strategies for choosing the optimal value of $K$.

@@

Of note, the MSN density can be expressed more compactly in terms of the matrix skew normal (MatSN) density (Chen and Gupta 2005). As we will see in Section 3.6, the matrix representation of the MSN distribution admits convenient conjugate prior distributions for the regression parameters and scale matrices, which in turn leads to efficient Gibbs sampling for posterior inference. Let $\mathbf{Y}_k$ be an ${n_k \times J}$ response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n {1}_{(z_i = k)}$ is the number of observations in cluster $k$. From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_k$ is distributed as
\begin{eqnarray*}
\mathbf{Y}_{k} &\sim& \text{MatSN}_{n_k \times J}(\mathbf{M}_k,\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)\\
%\text{vec}(\mathbf{Y}_k) =  (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})^T\\
\text{vec}(\mathbf{M}_k) &=& (\boldsymbol\zeta_{k1}^T,...,\boldsymbol\zeta_{kn_k}^T)^T,
\end{eqnarray*}
where $\boldsymbol\zeta_{ki} = \mathbf{X}_i \boldsymbol\beta_k$ as in equation (\ref{eq:msnreg}), $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})^T$, $\mathbf{I}_{n_k}$ is the $n_k\times n_k$ identity matrix, and $\boldsymbol\Omega_k$ is the $J\times J$ scale matrix defined above in equation (\ref{eq:msndens}). From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_{k}$, conditional on the $n_k\times 1$ vector of random effects $\mathbf{t}_{k} = (t_1,...,t_{n_k})^T$, is jointly distributed in matrix form as
\[
\mathbf{Y}_k | \mathbf{t}_k \sim \text{MatNorm}_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),
\]
where $\text{MatNorm}_{n_k \times J}(\cdot)$ denotes a $n_k\times J$ matrix normal density, $\text{vec}(\mathbf{M}_k) = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ is an $n_kJ\times 1$ mean vector, $\mathbf{X}_k$ is an $n_kJ \times Jp$ design matrix, $\boldsymbol\beta_k$ is the $(Jp)\times 1$ vector of regression coefficients defined in equation (\ref{eq:msnreg}), and $\boldsymbol\Sigma_k$ is the $J\times J$ conditional covariance of $\boldsymbol\epsilon_{ik}$ given in equation (\ref{eq:msnreg}).

\subsection{Multinomial Regression for the Cluster Indicators}
To accommodate heterogeneity in the cluster-membership probabilities, we model $\pi_{ik}$ as a function of covariates using a multinomial logit model 
\begin{eqnarray}
\pi_{ik} = \Pr(z_i = k|\mathbf{w}_i) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{h = 1}^K \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}},~ k=1,\ldots,K,
\end{eqnarray}
where $\mathbf{w}_i$ is an $r\times 1$ vector of subject-level covariates, $\boldsymbol\delta_k$ is a $r\times 1$ vector of regression parameters associated with membership in cluster $k$. For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$. Under this model, $z_i|\mathbf{w}_i \sim \text{Multinom}(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$. During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments as detailed later in section (\textbf{what section}). By allowing the cluster probabilities to vary across subjects, our model can be viewed as a \textit{mixture of experts} model in which $\pi_{ki}$ acts as a \textit{gating function} (\textbf{cite book}).

@@

By assuming a normal prior for $\boldsymbol\delta_k$, as detailed in Section 3.4, the full conditional distribution for $\boldsymbol\delta_k$ is normal with parameters that follow standard forms.

@@

To accommodate missing at random (MAR) responses, we propose a convenient imputation algorithm that can be implemented ``online'' as part of the Gibbs sampler. In Section 6, we discuss extensions to allow for non-ignorable missingness. Suppose $\mathbf{y}_i$ has $q_i\in (1,\ldots,J)$ observed values, denoted $\mathbf{y}^{obs}_i$, and $J-q_i$ intermittent missing values, denoted $\mathbf{y}^{miss}_i$. We can make use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate normal distribution given $(z_i,t_i,\mathbf{y}^{obs}_i)$:

@@

\boldsymbol\mu^{cond}_{ki}&=& \boldsymbol\mu^{miss}_{ki} + \boldsymbol\Sigma_{12} \boldsymbol\Sigma_{22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\mu^{obs}_{ki}) \label{eq:impute}\\
\boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{11} - \boldsymbol\Sigma_{12}\boldsymbol\Sigma_{22}^{-1}\boldsymbol\Sigma_{21},\nonumber \\

@@

$\boldsymbol\Sigma_{k11}$ is a $J-q_i \times J-q_i$ matrix containing the rows and columns of $\boldsymbol\Sigma_k$ corresponding to inicies of $\mathbf{y}_{i}$ where missingness occurs. Similarly, $\boldsymbol\Sigma_{k12}$ is a $J-q_i \times q_i$ matrix containing the rows of $\boldsymbol\Sigma_k$ that correspond to missing indices of $\mathbf{y}_i$, but columns of $\boldsymbol\Sigma_k$ that correspond to observed indices of $\mathbf{y}_i$. The remaining partitions $\boldsymbol\Sigma_{k21}, \ \text{and} \ \boldsymbol\Sigma_{k22}$ are defined in the same manner. Likewise, $\boldsymbol\mu_{ki} = \mathbf{X}_i \boldsymbol\beta_k$ is partitioned into $\boldsymbol\mu_{ki}^{miss}$ and $\boldsymbol\mu_{ki}^{obs}$ with respect to the missing and observed indices of $\mathbf{y}_i$, respectively. These results follow from conventional multivariate normal theory, which we detail in the Web Appendix. An attractive feature of this imputation algorithm is that it avoids multiplicative run-time scaling in $m$, the number of imputations (Gelman \textit{et al.} 2013; Zhou and Reiter, 2010). Our approach also provides more opportunities to explore the missing data parameter space than does multiple imputation, since missing values are drawn at each MCMC iteration, and often in practice $n_{sim} >> m$, where $n_{sim}$ is the number of MCMC iterations (\textbf{find a reference}). In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses improves inferences over available-case analysis.

@@

\subsubsection{Prior Specification}
We adopt a fully Bayesian inferential approach and assign prior distributions to all model parameters. Conveniently, all parameters admit conditionally conjugate priors, which greatly improves posterior computation via a data-augmented Gibbs sampler. For the MSN model component, we adopt a conditionally independent prior structure for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$, where $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k) = p(\boldsymbol\Sigma_k)p(\boldsymbol\beta_k|\boldsymbol\Sigma_k)$. We choose the normal-inverse-Wishart distribution for $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k)$ by specifying $\boldsymbol\Sigma_k \sim \text{IW}(\mathbf{V}_{0k},\nu_{0k})$ and $\boldsymbol\beta_k|\boldsymbol\Sigma_k \sim \text{N}_{Jp}(\mathbf{b}_k,\mathbf{I}_p \otimes \boldsymbol\Sigma_k)$. We assign the skewness coefficients $\boldsymbol\psi_k$ a conjugate $\text{N}_J(\mathbf{m}_k,\mathbf{P}_k)$ prior. In practice, the updates of $\boldsymbol\beta_k$ and $\boldsymbol\psi_k$ can be combined into one step by defining the $(Jp + J) \times 1$ vector $\boldsymbol\beta^*_k = (\boldsymbol\beta_k^T,\boldsymbol\psi_k^T)^T$ for which we assume a $\text{N}_{Jp+J}(\mathbf{b}^*_k,\mathbf{I}_{(p+1)} \otimes \boldsymbol\Sigma_k)$ prior.

To make use of the Matrix Normal representation introduced previously in Section (\textbf{what section}), we define the matrix of regression parameters $\mathbb{B}^*_k$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k$. We assign $\mathbb{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$ prior, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k}) = \mathbf{b}^*_k$, $\mathbf{I}_{p+1}$ is the identity matrix of dimensions $(p+1)\times(p+1)$, and $\boldsymbol\Sigma_k$ is the model error. This prior structure allows for convinient Matrix Normal updates of the MSN regression parameters as described in Section (\textbf{what section}).

@@

\subsubsection{Posterior Inference} 
The above prior specification induces closed-form full conditionals that can be efficiently updated as part of a Gibbs sampler outlined below. A programatic sketch of our MCMC algorithm is given in Table (\ref{alg:MCMC}). Additional details, including derivations can be found in the Web Appendix.

\paragraph{\textbf{Step 1}: Conditional MSN Imputation} The sampler begins with imputing missing values $\mathbf{y}_i^{miss}$ conditional on current simulated values of $z_i = k$ and $t_i$ as well as the associated $\mathbf{y}_i^{obs}$. This is accomplished by drawing $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in equation \ref{eq:impute} for all $i = 1,...,n$. Conditional imputation concludes with construction of a complete outcome vector $\mathbf{y}_i$ by merging the now imputed $\mathbf{y}_i^{miss}$ with $\mathbf{y}_i^{obs}$ with respect for the original order of $\mathbf{y}_i$.

@@

See Algorithm \ref{alg:MCMC}.

@@

A common challenge for Bayesian mixtures models is the so-called ``label switching" problem, in which draws of cluster-specific parameters may be assigned to the wrong cluster at some point during the MCMC simulation, rendering summaries of class-specific parameters incoherent. After conducting simulation studies under a wide variety of realistic parameter settings, as detailed in Section 4, we found little evidence of label switching in posterior draws of $\mathbf{z}$. When label switching was observed, we implemented \textit{post hoc} relabling algorithms as described in Papastamoulis 2016 and implemented in the \texttt{label.switching} package in \texttt{R}.

@@

We model fit diagnostics, we choose the WAIC statistic introduced by Watanabe (2010). To specify the number of cluster $K$, we follow Neelon and Chung (2016), who use WAIC to determine the optimal number of latent factors in a Bayesian latent factor zero-inflated Poisson model for the analysis of correlated and zero-inflated count data. In Section 6, we discuss possible extentions to other methods for choosing $K$.

@@

For cluster $k$, we can collect all observations $\mathbf{y}_i$ ($i = 1,...,n_k$) into the $n_k \times 4$ matrix $\mathbf{Y}_k$ and all covariates $\mathbf{X}_i$ into the $4n_k \times 8$ matrix $\mathbf{X}_k$ (\textbf{clash of notation here}) to utilize the matrix Normal model specification detailed in Section 3, which allows for simultaneous updates of all MSN regression parameters in cluster $k$. For the multinomial regression model component of this simulation, we model the class labels $z_i$ as a function of an intercept and one baseline covariate, thus $r = 2$. We leave the problem of missing data to simulation study \#2 and do not incorporate missing data into this simulation.

>>>>

\begin{abstract}
In studies of infant motor development, a crucial research goal is the identification of latent clusters of infants who experience delayed development, as this is a known risk factor for adverse outcomes later in life. However, there are a number of statistical challenges in modeling infant development: the data are typically skewed, exhibit intermittent missingness, and are highly correlated across the repeated measurements collected during infancy. Using data from the Nurture study, a cohort of over 600 mother-infant pairs followed from pregnancy to 12 months postpartum, we develop a flexible Bayesian finite mixture model for the analysis infant motor development. Our model has a number of attractive features. First, we adopt the multivariate skew normal distribution with cluster-specific parameters that accommodate the inherent correlation and skewness in the data. Second, we model the cluster membership probabilities using a novel application of the P\'olya-Gamma data-augmentation scheme, thereby improving predictions of the cluster membership allocations. Lastly, we impute missing responses under the missing at random assumption by drawing from appropriate conditional multivariate skew  normal distributions. Bayesian inference is achieved through straightforward Gibbs sampling, and can be implemented in available software such as \texttt{R}.  Through simulation studies, we show that the proposed model yields improved inferences over models that ignore skewness. In addition, our imputation method yields improvements compared to conventional missing data methods, including multiple imputation and complete or available case analysis. When applied to Nurture data, we identified two distinct development clusters: one characterized by delayed “U-shaped” development and a higher percentage of male infants and another characterized by more steady development and a lower percentage of males. The clusters also differed in terms of key demographic variables, such as infant race and maternal pre-pregnancy body mass index. These findings can aid investigators in targeting interventions during this critical early-life developmental window.
\end{abstract}

@@

Mixtures of multivariate non-symmetric distributions such as the multivariate skew-normal (MSN) distribution allow for the nuances of the marginal density to be captured with a more parsimonious set of mixture components. Mixtures of MSN distributions have been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010 and others), however in these models, focus lies primary on marginal density estimation, and inference on the mixture components (i.e. clusters) is often not discussed. More recently, the mixtures of skew-$t$ factor analysis (MSTFA) model has been proposed for settings in which cluster-specific inference is of primary interest (Lin \textit{et al.} 2018). However, an important feature not included in the MSTFA is the ability to explain individual-level cluster membership as a function of covariates of interest. Additionally, the parameter estimation procedure proposed by Lin et al. for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, including the ability to incorporate prior information into a model and make posterior probability statements. Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya-Gamma distribution.

@@

A ubiquitous feature of repeated measures studies is loss of data due to intermittent missingness and attrition. In the Bayesian setting, the standard approach to dealing with missing data is to perform multiple imputation, whereby $m$ imputed data sets are generated from a specified imputation model. After $m$ complete data sets are obtained, parameter estimates are combined across each data set to produce a final set of parameter estimates (Gelman \textit{et al.} 2013). This approach is not only computationally burdensome, requiring storage and analysis of an $m \times n_{rows} \times n_{cols}$ data array in addition to multiplication of total model run time by a factor of $m$, but it has been shown to produce unreliable inferences (Zhou and Reiter, 2010). We instead include an ``online" imputation imputation step in our Gibbs sampling procedure, whereby missing outcomes are updated at each iteration. This approach greatly increases the number of opportunities for exploration of the missing data parameter space and avoids the multiplication of total run time and number of parameters.

@@

The analysis of infant development trends in the Nurture data presents a number of statistical challenges that motivate our proposed model. First, as depicted in Figure 1, the residuals from repeated measures models of Bayley composite scores exhibit skewness even after adjusting for covariates such as race, sex, and birthweight. This suggests that the assumption of conditional normality made by standard repeated measures models is violated, and a distinguishing feature of the data, skewness, is not being accounted for.

@@

The Nurture data also feature intermittent missingness in Bayley composite scores throughout the study period. Of the total cohort (N = 666), 429 (64.4 \%) observations were available at three months, 435 (65.3 \%) observations were available at six months, 418 (62.8 \%) observations were available at nine months, and 437 (65.6 \%) observations were available at twelve months. As such, we require a modeling framework capable of dealing with missing data.

@@

A primary goal of the Nurture study is to identify clusters of infants characterized by distinct motor development trajectories throughout the first year of life. To address this aim, we propose a flexible finite mixture model that accommodates relevant features of the data, such as skewness and dependence among the responses. For $i = 1,...,n$, let $\mathbf{y}_{i}=(y_{i1},\ldots,y_{iJ})^T$ be a $J \times 1$ vector of responses (i.e., Baley composite scores) for subject $i$ across the $J$ measurement occasions. For the analysis of the Nurture data, we propose a finite mixture model of the form
\begin{equation}
f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ki} f(\mathbf{y}_i|\boldsymbol\theta_k),
\end{equation}
where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$) and $\pi_{ki}$ is a subject-specific mixing weight representing the probability that subject $i$ belongs to cluster $k$. For now we assume that $K$ is fixed; in Section 4, we discuss model selection strategies for choosing the optimal value of $K$.

@@

Of note, the MSN density can be expressed more compactly in terms of the matrix skew normal (MatSN) density (Chen and Gupta 2005). As we will see in Section 3.6, the matrix representation of the MSN distribution admits convenient conjugate prior distributions for the regression parameters and scale matrices, which in turn leads to efficient Gibbs sampling for posterior inference. Let $\mathbf{Y}_k$ be an ${n_k \times J}$ response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n {1}_{(z_i = k)}$ is the number of observations in cluster $k$. From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_k$ is distributed as
\begin{eqnarray*}
\mathbf{Y}_{k} &\sim& \text{MatSN}_{n_k \times J}(\mathbf{M}_k,\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)\\
%\text{vec}(\mathbf{Y}_k) =  (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})^T\\
\text{vec}(\mathbf{M}_k) &=& (\boldsymbol\zeta_{k1}^T,...,\boldsymbol\zeta_{kn_k}^T)^T,
\end{eqnarray*}
where $\boldsymbol\zeta_{ki} = \mathbf{X}_i \boldsymbol\beta_k$ as in equation (\ref{eq:msnreg}), $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})^T$, $\mathbf{I}_{n_k}$ is the $n_k\times n_k$ identity matrix, and $\boldsymbol\Omega_k$ is the $J\times J$ scale matrix defined above in equation (\ref{eq:msndens}). From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_{k}$, conditional on the $n_k\times 1$ vector of random effects $\mathbf{t}_{k} = (t_1,...,t_{n_k})^T$, is jointly distributed in matrix form as
\[
\mathbf{Y}_k | \mathbf{t}_k \sim \text{MatNorm}_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),
\]
where $\text{MatNorm}_{n_k \times J}(\cdot)$ denotes a $n_k\times J$ matrix normal density, $\text{vec}(\mathbf{M}_k) = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ is an $n_kJ\times 1$ mean vector, $\mathbf{X}_k$ is an $n_kJ \times Jp$ design matrix, $\boldsymbol\beta_k$ is the $Jp\times 1$ vector of regression coefficients defined in equation (\ref{eq:msnreg}), and $\boldsymbol\Sigma_k$ is the $J\times J$ conditional covariance of $\boldsymbol\epsilon_{ik}$ given in equation (\ref{eq:msnreg}).

\subsection{Multinomial Regression for the Cluster Indicators}
To accommodate heterogeneity in the cluster-membership probabilities, we model $\pi_{ik}$ as a function of covariates using a multinomial logit model 
\begin{eqnarray}
\pi_{ik} = \Pr(z_i = k|\mathbf{w}_i) = \frac{\text{e}^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{h = 1}^K \text{e}^{\mathbf{w}_i^T \boldsymbol\delta_{h}}},~ k=1,\ldots,K,
\end{eqnarray}
where $\mathbf{w}_i$ is an $r\times 1$ vector of subject-level covariates, $\boldsymbol\delta_k$ is a $r\times 1$ vector of regression parameters associated with membership in cluster $k$. For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$. Under this model, $z_i|\mathbf{w}_i \sim \text{Multinomial}(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$. During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments as detailed later in section (\textbf{what section}). By allowing the cluster probabilities to vary across subjects, our model can be viewed as a \textit{mixture of experts} model in which $\pi_{ki}$ acts as a \textit{gating function} (\textbf{cite book}).

@@

By assuming a normal prior for $\boldsymbol\delta_k$, as detailed in Section 3.4, the full conditional distribution for $\boldsymbol\delta_k$ is normal with parameters that follow standard forms, which we provide in Section (\textbf{what section}).

@@

To accommodate missing at random (MAR) responses, we propose a convenient imputation algorithm that can be implemented ``online'' as part of the Gibbs sampler. In Section 6, we discuss extensions to allow for non-ignorable missingness (i.e. observations missing not at random). Suppose $\mathbf{y}_i$ has $q_i\in (1,\ldots,J)$ observed values, denoted $\mathbf{y}^{obs}_i$, and $J-q_i$ intermittent missing values, denoted $\mathbf{y}^{miss}_i$. We can make use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate normal distribution given $(z_i,t_i,\mathbf{y}^{obs}_i)$:

@@

\boldsymbol\mu^{cond}_{ki}&=& \boldsymbol\mu^{miss}_{ki} + \boldsymbol\Sigma_{k12} \boldsymbol\Sigma_{k22}^{-1}(\mathbf{y}^{obs}_i - \boldsymbol\mu^{obs}_{ki}) \label{eq:impute}\\
\boldsymbol\Sigma^{cond}_k&=& \boldsymbol\Sigma_{k11} - \boldsymbol\Sigma_{k12}\boldsymbol\Sigma_{k22}^{-1}\boldsymbol\Sigma_{k21},\nonumber \\

@@

$\boldsymbol\Sigma_{k11}$ is a $(J-q_i) \times (J-q_i)$ matrix containing the rows and columns of $\boldsymbol\Sigma_k$ corresponding to inicies of $\mathbf{y}_{i}$ where missingness occurs. Similarly, $\boldsymbol\Sigma_{k12}$ is a $(J-q_i) \times q_i$ matrix containing the rows of $\boldsymbol\Sigma_k$ that correspond to missing indices of $\mathbf{y}_i$, but columns of $\boldsymbol\Sigma_k$ that correspond to observed indices of $\mathbf{y}_i$. The remaining partitions $\boldsymbol\Sigma_{k21}, \ \text{and} \ \boldsymbol\Sigma_{k22}$ are defined in the same manner. Likewise, $\boldsymbol\mu_{ki} = \mathbf{X}_i \boldsymbol\beta_k$ is partitioned into $\boldsymbol\mu_{ki}^{miss}$ and $\boldsymbol\mu_{ki}^{obs}$ with respect to the missing and observed indices of $\mathbf{y}_i$, respectively. These results follow from conventional multivariate normal theory, which we detail in the Web Appendix. An attractive feature of this imputation algorithm is that it avoids multiplicative run-time scaling in $m$, the number of imputations (Gelman \textit{et al.} 2013; Zhou and Reiter, 2010). Our approach also provides more opportunities to explore the missing data parameter space than does multiple imputation, since missing values are drawn at each MCMC iteration, and often in practice $n_{sim} >> m$, where $n_{sim}$ is the number of MCMC iterations (\textbf{find a reference}). In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses improves inferences over available-case analysis.

@@

\subsubsection{Prior Specification}
We adopt a fully Bayesian inferential approach and assign prior distributions to all model parameters. Conveniently, all parameters admit conditionally conjugate priors, which greatly improves posterior computation via a data-augmented Gibbs sampler. For the MSN model component, we adopt a conditionally independent prior structure for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$, where $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k) = p(\boldsymbol\Sigma_k)p(\boldsymbol\beta_k|\boldsymbol\Sigma_k)$. We choose the normal-inverse-Wishart distribution for $p(\boldsymbol\beta_k,\boldsymbol\Sigma_k)$ by specifying $\boldsymbol\Sigma_k \sim \text{IW}(\mathbf{V}_{0k},\nu_{0k})$ and $\boldsymbol\beta_k|\boldsymbol\Sigma_k \sim \text{N}_{Jp}(\mathbf{b}_k,\mathbf{I}_p \otimes \boldsymbol\Sigma_k)$. We assign the skewness coefficients $\boldsymbol\psi_k$ a conjugate $\text{N}_J(\mathbf{m}_k,\mathbf{P}_k)$ prior. In practice, the updates of $\boldsymbol\beta_k$ and $\boldsymbol\psi_k$ can be combined into one step by defining the $(Jp + J) \times 1$ vector $\boldsymbol\beta^*_k = (\boldsymbol\beta_k^T,\boldsymbol\psi_k^T)^T$ for which we assume a $\text{N}_{Jp+J}(\mathbf{b}^*_k,\mathbf{I}_{(p+1)} \otimes \boldsymbol\Sigma_k)$ prior, where $\mathbf{b}^*_k$ is formed by concatenating $\mathbf{b}_k$ and $\mathbf{m}_k$.

To make use of the Matrix Normal representation introduced previously in Section (\textbf{what section}), we define the matrix of regression parameters $\mathbb{B}^*_k$, where $\text{vec}(\mathbb{B}^*_k) = \boldsymbol\beta^*_k$. We assign $\mathbb{B}^*_k|\boldsymbol\Sigma_k$ a $\text{MatNorm}(\mathbf{B}^*_{0k},\mathbf{I}_{p+1},\boldsymbol\Sigma_k)$ prior, where $\mathbf{B}^*_{0k}$ is a matrix of location parameters such that $\text{vec}(\mathbf{B}^*_{0k}) = \mathbf{b}^*_k$, $\mathbf{I}_{p+1}$ is the identity matrix of dimensions $(p+1)\times(p+1)$, and $\boldsymbol\Sigma_k$ controls the correlation of the model error. This prior structure allows for convinient Matrix Normal updates of the MSN regression parameters as described in Section (\textbf{what section}).

@@

\subsubsection{Posterior Inference} 
The above prior specification induces closed-form full conditionals that can be efficiently updated as part of a Gibbs sampler outlined below. A programatic sketch of our MCMC algorithm is given in Table (\ref{alg:MCMC}). Additional details, including derivations of full conditionals and MCMC diagnostics can be found in the Web Appendix.

\paragraph{\textbf{Step 1}: Conditional MSN Imputation} The sampler begins with imputing missing values $\mathbf{y}_i^{miss}$ conditional on current  values of $z_i = k$ and $t_i$ as well as the associated $\mathbf{y}_i^{obs}$ observed data vector. This is accomplished by drawing $\mathbf{y}_i^{miss}$ from $\text{N}_{J-q_i}(\boldsymbol\mu^{cond}_{ki},\boldsymbol\Sigma^{cond}_k)$ as described in equation \ref{eq:impute} for all $i = 1,...,n$. Conditional imputation concludes with construction of a complete outcome vector $\mathbf{y}_i$ by merging the now imputed $\mathbf{y}_i^{miss}$ with $\mathbf{y}_i^{obs}$ with respect for the original ordering of the elements of $\mathbf{y}_i$.

@@

See Table 1.

@@

A common challenge for Bayesian mixture models is the so-called ``label switching" problem, in which draws of cluster-specific parameters may be assigned to the wrong cluster at some point during the MCMC simulation, rendering summaries of class-specific parameters incoherent. After conducting simulation studies under a wide variety of realistic parameter settings, as detailed in Section 4, we found little evidence of label switching in posterior draws of $\mathbf{z}$. When label switching was observed, we implemented \textit{post hoc} relabling algorithms as described in Papastamoulis 2016 and implemented in the \texttt{label.switching} package in \texttt{R}.

@@

For model fit diagnostics, we choose the WAIC statistic introduced by Watanabe (2010). To specify the number of cluster $K$, we follow Neelon and Chung (2016), who use WAIC to determine the optimal number of latent factors in a Bayesian latent factor zero-inflated Poisson model for the analysis of correlated and zero-inflated count data. In Section 6, we discuss possible extentions to other methods for choosing $K$.

@@

For cluster $k$, we can collect all observations $\mathbf{y}_i$ ($i = 1,...,n_k$) into the $n_k \times 4$ matrix $\mathbf{Y}_k$ and all covariates $\mathbf{x}_i$ into the $4n_k \times 8$ matrix $\mathbf{X}_k$ to utilize the matrix Normal model specification detailed in Section 3, which allows for simultaneous updates of all MSN regression parameters in cluster $k$. For the multinomial regression model component of this simulation, we model the class labels $z_i$ as a function of an intercept and one baseline covariate, thus $r = 2$. We leave the problem of missing data to simulation study \#2 and do not incorporate missing data into this simulation.
